Fast Automatic Bayesian Cubature Using Lattice Sampling

### 2019-07-20 ResponseToReferees.tex

\newcommand{\vV}{\boldsymbol{V}}
\newcommand{\vone}{\boldsymbol{1}}

\begin{document}
\title[Response for Fast Automatic Bayesian Cubature]{Response to Referees of the manuscript \\
\emph{Fast Automatic Bayesian Cubature Using Lattice Sampling}}
\maketitle

\newcommand{\Response}[1]{{\color{blue}#1}}
\renewcommand{\vec}[1]{\boldsymbol{#1}}
\newcommand{\vtheta}{\vec{\theta}}

Thanks to the the referees and the editors for their very helpful comments.  We have revised our manuscript substantially in response. \Response{Specific concerns are addressed point-by-point below in blue.}. We hope that all will find the revision satisfactory.

@@

\item Section 5.1. The Baker's transform is introduced first and then the authors talk about
more general transforms without saying so. For the Baker's transform the
substitution written with $\Psi'$ is undefined. This paragraph should be
restructured. \Response{Done.}

\end{itemize}

\section*{Reviewer 2: The authors study a stopping criterion \ldots}

There is no theory available that covers the whole process (estimation and cubature with credible interval); instead, the authors present numerical experiments to demonstrate the performance of three variants of their new algorithm. \Response{We would disagree with the sweeping nature of the claim that ``there is no theory''.  Our method is taking the established theory of Bayesian numerical analysis and using it to develop adaptive stopping criteria.  In our revised manuscript we summarize our results in three theorems.  We recognize some gaps in the theory.}

@@

\item Sec.\ 3.2--3.4 Does the operation count also include the computational cost for the minimization problem to estimate $\theta$? I suspect that this is not the case, since there is no assumption concerning the dependence of $C_\theta$ on $\theta$. The same question arises in the setting of Section 4.  \Response{You are correct.  This has been clarified.}

\item Sec.\ 4.1 and 4.2 As the authors state, larger $r$ implies a greater degree of smoothness of the kernel. This should lead to a greater amount of smoothness of the integrands, and thus to a fast convergence of suitable cubature formulas. It might be that this is not exploited by the integration lattices (plus periodization), as studied in the present paper. Please comment.  \Response{Greater smoothness of the covariance kernel implies an \emph{assumption} of greater smoothness of the integrands, which may or may not be justified.  Lattice rules work well for all degrees of smoothness. This point is mentioned at the end of Section 4.2.}

\item p.\ 9, eqn. (37) $i, j = 1$. \Response{Done.}

\item p.\ 9, bottom, right Assumption (25c) \ldots \Response{The rewriting of this section has eliminated that sentence.}

@@

\item p.\ 12, center left  \ldots differs depends \ldots ?  \Response{Fixed.}

\item Sec.\ 5.1 Is there some kind of a universal periodization that ‘works’ for every value of $r$?  \Response{One could choose an infinitely smooth periodization, but this would change the integrand to potentially make its covariance $s$ large. This is mentioned at the end of Section 5.1.}

\item Sec. 5.2 You take $r = 2$ or $r = 1$ for the periodization in these examples. Does this mean that $\theta = (2,\gamma)$ is considered for the shift-invariant kernel, or do you still consider $\theta = (r,\gamma)$ with $r \in \mathbb{N}$ as the unknown parameter?  \Response{For technical simplicity we fix $r$.  We mention that in Section 5.2.}

\item I suggest not to use $r$ also in the drift term in the option pricing example. \Response{Excellent suggestion.  Replaced by $R$.}

@@

\item Two lines before Equation (20): ``confidence interval" and not ``credibility interval" like the other ones?  \Response{Fixed.}

\item Five lines after (21): err$_{\rm CI}$ refers to what exactly? The value in (20)?  Clarify.  \Response{Re-written to clarify.}

\item In (24), are $C_1,\dots,C_n$ the columns of C?  \Response{Yes.}

\item It took me a while to figure out what the four different colors of the points in Figures 5 to 13 mean.  There is a continuum of colors on the right, but only four specific colors for the points are used, so I think it would be useful to say explicitly what each of these four colors  represent.  In the text on page 12, it would be useful to say for example that $\epsilon = 10^{-5}$ corresponds to the green points.   By the way, are there 100 green points, for the 100 independent shifts?  \Response{We have modified our experiments to randomly choose from a continuum of error tolerances.  We have also explained that each experiment correspondes to a different random shift.}

@@

\item Section 5.2, line 8: correct ``differs depends". \Response{Fixed.}

\item Page 14: The three figures look all the same!  Any comment on that?  \Response{They are similar, but not the same.}

\item In this example, does it really make sense to assume that the integrand comes from a Gaussian process?  \Response{Theoretically, perhaps not, but with a nonzero mean, maybe this is not too much of a stretch.  Bayesian cubature assuming non-Gaussian processes is much more challenging.  We have addressed this at the end of Section 5.2}

\item Four lines below Figure 13:  "A noticeable aspect from the plots is how much ....":  It may be good to explain exactly how we see that in the plots. \Response{We have elaborated more.}

@@

\item Reference [1]:  "Griolami"?   Mark may not recognize himself!  \Response{Sorry.  Fixed.}

\end{itemize}

\section*{Editor: We have received two very careful reviews of this paper \ldots}

The revision should at a minimum point out the connection to the FFT. \Response{Done.}

Because the goal is to reach people outside of PN some additional explanations are in order. \Response{We have tried.}

\begin{enumerate}
\renewcommand{\labelenumi}{\arabic{enumi}.}
    \item The results in Section 2.2.2 were confusing. \ldots \Response{This has been re-written}

@@

    \item It is weird to use complex numbers for the eigenvectors of a real symmetric matrix \ldots \Response{While it may be strange, it is convenient since we are working with FFT's of real vectors, which are complex vectors.  Some explanation is provided at the beginning of Section 3.1.}

\item In(24), why is $V^H =nV^{-1}$ and not just $V^{-1}$? \ldots \Response{This normalization allows $\vV_1 = \vone$.}

\item The argument at the end of section 4.3 involving is very slippery. \ldots \Response{Section 4 has been substantially re-written.}

### 2019-07-20 BayesianCubaturePaper2018 - Rev1b.tex
revising paper

@@

Assumptions \eqref{fastcompAssumpA} and \eqref{fastcompAssumpB} can be verified by \eqref{PermMat}, \eqref{Keig}, and \eqref{Clateig}.  Assumption \eqref{fastcompAssumpC} is satisfied because $\mV^H \vb = \mW^H \mP^T \vb$ is just the permutation of the rows of the discrete Fourier transform of a vector.  This can be performed in $\Order(n \log(n))$ operations by the FFT algorithm. A proper scaling of the kernel $K_\vtheta$ ensures that assumption \eqref{addAssump} is satisfied.

@@

Periodizing variable transforms are used in the numerical examples below. In some cases, they can speed the convergence of the Bayesian cubature because they allow one to take advantage of smoother kernels.

@@

Three integrals were evaluated using the GAIL algorithm \code{cubBayesLattice\_g}:  a multivariate normal probability, the Keister's example, and an option pricing example.  
The sequences $\{\vx_i\}_{i=1}^\infty$ were the randomly shifted lattice node sequences supplied by GAIL. 
For each integral,  each of our stopping criteria---empirical Bayes, full Bayes, and generalized cross-validation---our algorithm was run for $400$ different randomly chosen error tolerances. The error tolerances were chosen randomly in an interval depending on the difficulty of the problem. In each run, the nodes were also randomly shifted. The accuracy of the algorithm differs based on the shift.
For each test, the execution times were plotted against $\abs{\mu - \hmu}/\varepsilon$.  We expect $\abs{\mu - \hmu}/\varepsilon$ to be no greater than one, but hope that it is not too much smaller than one, which would indicate a stopping criterion that is too conservative.

@@

The price of financial derivatives can often be modeled by high dimensional integrals. If the underlying asset is described in terms of a discretized geometric Brownian motion, then the fair price of the option is:
\begin{equation*}
\mu = \int_{\reals^d} \text{payoff}(\vz) \frac{\exp(\frac 12 \vz^T\mSigma^{-1}\vz)}{\sqrt{(2\pi)^d \det(\mSigma)}} \, \dvz = \int_{[0,1]^d} f(\vx) \, \dvx,
\end{equation*} 
where {payoff($\cdot$)} defines the discounted payoff of the option,
\begin{align*}
\mSigma &= (T/d) \bigl(\min(j,k) \bigr)_{j,k=1}^d = \mL \mL^T,\\
f(\vx) &= \text{payoff} \left(\mL 
\begin{pmatrix}
\Phi^{-1}(x_1) \\ \vdots \\ \Phi^{-1}(x_d)
\end{pmatrix} \right).
\end{align*}
The Asian arithmetic mean call option has a payoff of the form
\begin{align*}
\text{payoff}(\vz) &= \max\left( \frac 1d  \sum_{j=1}^d S_j(\vz) - K, 0 \right) \me^{-\upsilon T}, \\
\text{where}\;
S_j(\vz) &= S_0 \exp\bigl((\upsilon-\sigma^2/2)jT/d + \sigma \sqrt{T/d} z_j \bigr).
\end{align*}
Here, $T$ denotes the time to maturity of the option, $d$ the number of time steps, $S_0$ the initial price of the stock, $\upsilon$ the interest rate, $\sigma$ the volatility, and $K$ the strike price.

Figures \ref{fig:optprice-guaranteed-MLE}, \ref{fig:optprice-guaranteed-FB} and 
\ref{fig:optprice-guaranteed-GCV} summarize the numerical results for this example using
$
T = 1/4, \ \ d = 13, \ \ S_0 = 100, \ \ \upsilon =  0.05, \ \ \sigma = 0.5, \ \ K = 100.
$
Moreover, $\mL$ is chosen to be the matrix of eigenvectors of $\mSigma$ times the square root of the diagonal matrix of eigenvalues of $\mSigma$.
Because the integrand has a kink caused by the $\max$ function, it does not help to use a periodizing transform that is very smooth.  We choose the Baker's transform \eqref{eq:bakerTrans} and $r = 1$.

@@

A noticeable aspect from the plots is how much the error bounds differ from the true error. For option pricing example, the error bound is not as conservative as it is for the multivariate normal and Keister examples. A possible reason is that the latter integrands are significantly smoother than the covariance kernel assumed.  This is a matter for further investigation.

\iffalse
\subsection{Diagnostics for the Gaussian Process Assumption}

The starting point for our Bayesian cubature is the assumption that the integrand arises from a Gaussian process. This means that the function data, $\vf$ satisfy a multivariate Gaussian distribution, as in \eqref{eqn:fGaussDist}.  The transformed data, $\vZ = ( n\mLambda)^{-\frac 12} \mV^H(\vf - m \vone)$ has zero mean and is also uncorrelated because
\begin{align*}
\cov (\vZ) 
&= \frac 1n \Ex\left[  
\mLambda^{-\frac 12} \mV^H (\vf - m \vone)
(\vf - m \vone)^T \mV \mLambda^{-\frac 12}
\right]
\\
&=
 \mLambda^{-\frac 12} \mV^H 
\Ex\left[ (\vf - m \vone)
(\vf - m \vone)^T \right] \mV \mLambda^{-\frac 12}
\\
&=
\frac{1}{n} \mLambda^{-\frac 12} \mV^H 
\frac 1n \mV \mLambda \mV^H \mV \mLambda^{-\frac 12}
 = \mathsf{I}
\end{align*}
Thus, the elements of $\vZ$ are IID standard Gaussian random variables.

\figref{fig:mvn-normplot} and \figref{fig:keister-normplot} are normal probability plots of the $Z_i$ using empirical Bayes estimates of $m$ and $\vtheta$. \textbf{more goes here}.

\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{"figures/arbMean/Keister/C1sin/Keister Normplot d_2 bernoulli_2 Period_C1sin n_32768"}
	\caption{Normal plot : Keister function with arbMean assumption}
	\label{fig:keister-normplot}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{"figures/arbMean/MVN/C1sin/MVN Normplot d_2 bernoulli_2 Period_C1sin n_32768"}
	\caption{Normal plot : MVN with arbMean assumption}
	\label{fig:mvn-normplot}
\end{figure}

>>>>

Integration lattice cubature rules are known to have convergence rates that depend on the smoothness of the integrands \cite{DicEtal14a}.  Thus, we expect integration lattice node sets to perform well regardless of the smoothness of the covariance kernel.  The bigger concern is whether the derivatives of the integrand are as smooth as the covariance kernel implies.  This topic is touched upon again in Section  \ref{period_var_tx}.

@@

Assumptions \eqref{fastcompAssumpA} and \eqref{fastcompAssumpB} can be verified by \eqref{PermMat}, \eqref{Keig}, and \eqref{Clateig}.  Assumption \eqref{fastcompAssumpC} is satisfied because $\mV^H \vb = \mW^H \mP^T \vb$ is just the discrete Fourier transform of a vectors whose rows have been permuted.  This can be performed in $\Order(n \log(n))$ operations by the FFT algorithm. A proper scaling of the kernel $K_\vtheta$ ensures that assumption \eqref{addAssump} is satisfied.

@@

Periodizing variable transforms are used in the numerical examples below. In some cases, they can speed the convergence of the Bayesian cubature because they allow one to take advantage of smoother kernels. 
However, there is a trade-off.  Smoother periodizing transformations tend to give integrands $f$ with larger inferred $s$ values and thus wider credible intervals.

@@

Three integrals were evaluated using the GAIL algorithm \code{cubBayesLattice\_g}:  a multivariate normal probability, the Keister's example, and an option pricing example.  
The sequences $\{\vx_i\}_{i=1}^\infty$ are the randomly shifted lattice node sequences supplied by GAIL. 
For each integral,  each of our stopping criteria---empirical Bayes, full Bayes, and generalized cross-validation---our algorithm was run for $400$ different randomly chosen error tolerances. The error tolerances were chosen randomly in an interval depending on the difficulty of the problem. In each run, the nodes were also randomly shifted. The accuracy of the algorithm depends mildly on the shift; there is no universally optimal shift.
For each test, the execution times are plotted against $\abs{\mu - \hmu}/\varepsilon$.  We expect $\abs{\mu - \hmu}/\varepsilon$ to be no greater than one, but hopefully not too much smaller than one, which would indicate a stopping criterion that is too conservative.

Ideally, we would optimize both $r$ and $\eta$ simultaneously in the definition of our $C_\vtheta$ in \eqref{the_kernel_eqn_bernoulli}.  However, in these examples we fix $r$ and optimize $\eta$ only.  This is a technical challenge, not a limitation of our theory.

@@

The price of financial derivatives can often be modeled by high dimensional integrals. If the underlying asset is described in terms of a discretized geometric Brownian motion, then the fair price of the option is:
\begin{equation*}
\mu = \int_{\reals^d} \text{payoff}(\vz) \frac{\exp(\frac 12 \vz^T\mSigma^{-1}\vz)}{\sqrt{(2\pi)^d \det(\mSigma)}} \, \dvz = \int_{[0,1]^d} f(\vx) \, \dvx,
\end{equation*} 
where {payoff($\cdot$)} defines the discounted payoff of the option,
\begin{align*}
\mSigma &= (T/d) \bigl(\min(j,k) \bigr)_{j,k=1}^d = \mL \mL^T,\\
f(\vx) &= \text{payoff} \left(\mL 
\begin{pmatrix}
\Phi^{-1}(x_1) \\ \vdots \\ \Phi^{-1}(x_d)
\end{pmatrix} \right).
\end{align*}
The Asian arithmetic mean call option has a payoff of the form
\begin{align*}
\text{payoff}(\vz) &= \max\left( \frac 1d  \sum_{j=1}^d S_j(\vz) - K, 0 \right) \me^{-R T}, \\
\text{where}\;
S_j(\vz) &= S_0 \exp\bigl((R-\sigma^2/2)jT/d + \sigma \sqrt{T/d} z_j \bigr).
\end{align*}
Here, $T$ denotes the time to maturity of the option, $d$ the number of time steps, $S_0$ the initial price of the stock, $R$ the interest rate, $\sigma$ the volatility, and $K$ the strike price.

Figures \ref{fig:optprice-guaranteed-MLE}, \ref{fig:optprice-guaranteed-FB} and 
\ref{fig:optprice-guaranteed-GCV} summarize the numerical results for this example using
$
T = 1/4, \ \ d = 13, \ \ S_0 = 100, \ \ R =  0.05, \ \ \sigma = 0.5, \ \ K = 100.
$
Moreover, $\mL$ is chosen to be the matrix of eigenvectors of $\mSigma$ times the square root of the diagonal matrix of eigenvalues of $\mSigma$.
Because the integrand has a kink caused by the $\max$ function, it does not help to use a periodizing transform that is very smooth.  We choose the Baker's transform \eqref{eq:bakerTrans} and $r = 1$.

@@

One may question whether an integrand with non-negative values is well represented by a Gaussian process.  Since we allow a nonzero mean, this assumption is somewhat more palatable. Bayesian algorithms assuming non-Gaussian processes are more difficult, but this is an ares for further research.

A noticeable observation from the plots in all three examples is how much the credible interval widths differ from the true error. The credible interval widths are not much smaller than $\varepsilon$.  The ratio of the true error to $\varepsilon$ ranges from around one to as small as $10^{-7}$.  In the latter case, the credible interval widths are quite conservative.  For option pricing example, this less serious of an issue than for the multivariate normal and Keister examples. A possible reason is that the latter integrands are significantly smoother than the covariance kernel assumed.  This is a matter for further investigation.
